Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-01-11 paper/main.tex
fix some references to supp figs

(Figs.~\ref{fig:question-correlations},~\ldaVsBERT). The time series plot of
individual questions' correlations are different from each other when computed
using LDA (e.g., the traces can be clearly visually separated), whereas the
correlations computed from BERT embeddings of different questions all look very
similar. This tells us that LDA is capturing some differences in content

@@

At the opposite end of the spectrum from large language models, one could also
imagine \textit{simplifying} some aspects of our LDA-based approach by
computing simple word overlap metrics. For example, the Jaccard similarity
between text $A$ and $B$ is computed as the number of unique words in the
intersection of words from $A$ and $B$ divided by the number of unique words in
the union of words from $A$ and $B$. In a supplementary analysis (Supp.
Fig.~\jaccard), we compared the LDA-based question-lecture matches we reported
in Figure~\ref{fig:question-correlations} with the Jaccard similarities between
each question and each sliding window of text from the corresponding lecture.
As shown in Supplementary Figure~\jaccard, this simple word-matching approach
does not appear to capture the same level of specificity as the LDA-based
approach. Whereas the LDA-based approach often yields a clear peak in the
time series of correlations between each question and the corresponding lecture,
the Jaccard similarity-based approach does not. Furthermore, these LDA-based
matches appear to capture conceptual overlaps between the questions and
lectures (Supp.~Tab.~\matchTab), whereas simple word matching does not. For
example, one of the example questions examined in Supplementary
Figure~\jaccard~asks ``Which of the following occurs as a cloud of atoms gets
more dense?'' The LDA-based matches identify lecture timepoints where the
relevant \textit{topics} are discussed (e.g., when words like ``cloud,''
``atom,'' ``dense,'' etc., are mentioned \textit{together}). The Jaccard
similarity-based matches, on the other hand, are strong when \textit{any} of
these words are mentioned, even if they do not occur together.

>>>>

(Fig.~\ref{fig:question-correlations}, Supp.~Fig.~\ldaVsBERT). The time series
plot of individual questions' correlations are different from each other when
computed using LDA (e.g., the traces can be clearly visually separated), whereas
the correlations computed from BERT embeddings of different questions all look
very similar. This tells us that LDA is capturing some differences in content

@@

At the opposite end of the spectrum from large language models, one could also
imagine \textit{simplifying} some aspects of our LDA-based approach by
computing simple word overlap metrics. For example, the Jaccard similarity
between text $A$ and $B$ is computed as the number of unique words in the
intersection of words from $A$ and $B$ divided by the number of unique words in
the union of words from $A$ and $B$. In a supplementary analysis
(Supp.~Fig.~\jaccard), we compared the LDA-based question-lecture matches we
reported in Figure~\ref{fig:question-correlations} with the Jaccard similarities
between each question and each sliding window of text from the corresponding
lecture. As shown in Supplementary Figure~\jaccard, this simple word-matching
approach does not appear to capture the same level of specificity as the
LDA-based approach. Whereas the LDA-based approach often yields a clear peak in
the time series of correlations between each question and the corresponding
lecture, the Jaccard similarity-based approach does not. Furthermore, these
LDA-based matches appear to capture conceptual overlaps between the questions
and lectures (Supp.~Tab.~\matchTab), whereas simple word matching does not. For
example, one of the example questions examined in Supplementary
Figure~\jaccard~asks ``Which of the following occurs as a cloud of atoms gets
more dense?'' The LDA-based matches identify lecture timepoints where the
relevant \textit{topics} are discussed (e.g., when words like ``cloud,''
``atom,'' ``dense,'' etc., are mentioned \textit{together}). The Jaccard
similarity-based matches, on the other hand, are strong when \textit{any} of
these words are mentioned, even if they do not occur together.
